# Reproducible Research: Peer Assessment 1
The following packages must be loaded into R
```{r, echo=TRUE}
library(lubridate)    ## For dates
library(ggplot2)      ## For plottingt
library(dplyr)        ## Used by group_by and %>%
```

## Loading and preprocessing the data
It is assumed that the reader has set the correct working directory in R by function setwd()
```{r, echo=TRUE}
df <- read.csv("activity.csv", header = TRUE, sep = ',')
```
A first view and knowledge of the data is obtained by function str() and head()
```{r, echo=TRUE}
str(df)
'data.frame':	17568 obs. of  3 variables:
 $ steps   : int  NA NA NA NA NA NA NA NA NA NA ...
 $ date    : Factor w/ 61 levels "2012-10-01","2012-10-02",..: 1 1 1 1 1 1 1 1 1 1 ...
 $ interval: int  0 5 10 15 20 25 30 35 40 45 ...
head(df)
  steps       date interval
1    NA 2012-10-01        0
2    NA 2012-10-01        5
3    NA 2012-10-01       10
4    NA 2012-10-01       15
5    NA 2012-10-01       20
6    NA 2012-10-01       25
```
Convert the date from a factor to a date format and remove all rows from data set with step == NA  
and store the result in new object df_NoNA
```{r, echo=TRUE}
df$date <- ymd(df$date)
df_NoNA<-na.omit(df)
str(df_NoNA)
'data.frame':	15264 obs. of  3 variables:
 $ steps   : int  0 0 0 0 0 0 0 0 0 0 ...
 $ date    : POSIXct, format: "2012-10-02" "2012-10-02" "2012-10-02" "2012-10-02" ...
 $ interval: int  0 5 10 15 20 25 30 35 40 45 ...
 - attr(*, "na.action")=Class 'omit'  Named int [1:2304] 1 2 3 4 5 6 7 8 9 10 ...
  .. ..- attr(*, "names")= chr [1:2304] "1" "2" "3" "4" ...
```
2304 rows are omitted due to NA in df  
15264 rows in data frame df_NoNA

```{r, echo=TRUE}
head(df_NoNA)
    steps       date interval
289     0 2012-10-02        0
290     0 2012-10-02        5
291     0 2012-10-02       10
292     0 2012-10-02       15
293     0 2012-10-02       20
294     0 2012-10-02       25
```
## What is mean total number of steps taken per day?
Object StepsPerDay gives number of steps per day by grouping df_NoNA by date and summarize  
the steps
```{r, echo=TRUE}
StepsPerDay<- df_NoNA %>% group_by(date) %>% summarise(step=sum(steps))
head(StepsPerDay)
Source: local data frame [6 x 2]
        date  step
      (time) (int)
1 2012-10-02   126
2 2012-10-03 11352
3 2012-10-04 12116
4 2012-10-05 13294
5 2012-10-06 15420
6 2012-10-07 11015
```
A histogram gives the distribution of steps per day
```{r, echo=TRUE}
HistDate<-ggplot(StepsPerDay, aes(date, step)) + geom_bar(stat = "identity", colour = "steelblue", fill = "steelblue")
+ labs(title = "Total Number of Steps Per Day", x = "Date", y = "# Steps")  
HistDate
```
Plot here....  

The mean total of number of steps per day is 10766.19  
```{r, echo=TRUE}
mean(StepsPerDay$step)
[1] 10766.19
```
The median of the total number of steps per day is given by the 27th observation (half of the total # of observations)  
when the data is ordered by StepsPerDay and the value is 10765  
```{r, echo=TRUE}
median(StepsPerDay$step)
[1] 10765
```

## What is the average daily activity pattern?




## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
